Making Sequential Programs Deterministic

### 2019-03-22 design.tex
bullet pointing dataflow analysis and defaulting subsections.

\subsection{Dataflow analysis}

\subsection{Type Refinement}

\subsection{Discussion on defaulting rules}
Det - default, but @Poly is the default for parameters and return types
Choice of defaults on class type, constructors and method params and return types
Defaults on hashsets, hashmap, etc

\subsection{Avoid aliasing unsoundness}
Recall CLIMB-to-top rule and defaulting choice for local collections (to avoid too many warnings)
Required user annotations for collections and arrays

\subsection{Additional hardcoded rules for precision}
Example: equals on HashSet
Discussion on System properties - env, line, file and path separator.

@@

\subsection{About generics and upper bounds?}

>>>>

\subsection{Dataflow analysis and type refinement}
The checker framework performs automatic type inference for local types within method bodies.
It also performs dataflow analysis and refines the types of these locals, thereby reducing the annotation burden 
on the programmer.

\begin{itemize}
	\item Example code with explicit annotations on local variable, showing flow of values into this local variable
	\item Explain that CF refines the type of this local variable; inbuild dataflow analysis
	\item Constrast this with a type on a method argument that won't be type refined.
	\item CF allows type system developers to define additional refinement rules.
	\item Determinism checker performs type refinement for the receivers of sort() methods in Collection classes.
	\item Example client code where this type refinement results in precise results
\end{itemize}

\subsubsection{CLIMB-to-top and collection, array locals}

\begin{itemize}
	\item Every type system has a default annotation - the type that will be applied to unannotated locations.
	\item CF defaults types at CLIMB locations to top and type refines them. (steal some text from manual).
	\item CLIMB locations do not include type parameters. So this is problematic for determinism checker
	\item Example: Unannotated collection; its default would be @NonDet<@Det> which is invalid.
	\item So we default locals collections to @NonDet<@NonDet>. 
	\item This is sound but could produce a lot of FPs. Example assignment: @NonDet<@NonDet> collection = collection1.
	Type of collection1 is @Det<@Det> and it is a method argument.
	\item To avoid this, we require users to explicitly annotate local collections.
	\item Similarly for local arrays.
\end{itemize}

\subsection{Discussion on defaulting rules}

\begin{itemize}
	\item Determinism checker has @Det as the usual default. Reason being we expect the program to be 
	deterministic unless specified by programmer or nondeterminism from libraries that we have annotated (Random, Collections, etc).
	\item All method parameters and return types (including constructors) are @PolyDet.
	\item Reason: We want to be able to report nondeterministic behavior only where it is invalid.
	\item Example of a method with params and return types annotated as @PolyDet
	This method called from 2 places - one harmless(@Det instantiations), and another that call with @NonDet args and prints the retun type.
	\item Exception to this defaulting is the construction of HashSets and HashMaps because they can't be @Det.
	We explicitly annotate the return type on constructors as @OrderNonDet and disallow constructing @Det HashSet or HashMap.
\end{itemize}

\subsection{Additional hardcoded rules in the determinism checker  for precision}
Example: equals on HashSet
Discussion on System properties - env, line, file and path separator.

@@
